\documentclass{article}

% arXiv-bound preprint. Switch to \usepackage{neurips_2024} for anonymous
% submission, or \usepackage[final]{neurips_2024} for camera-ready.
\usepackage[preprint]{neurips_2024}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{graphicx}
\graphicspath{{../shared/figures/}{../experiments/analysis/figures/}}

\title{The Biology of Agentic Research Engineering: \\
       An Agent-Driven Variant Effect Prediction Workshop}

\author{%
  C\'esar M. Valdez C\'ordova \\
  Latent Reasoning Works \\
  \texttt{cesar@latent-reasoning.works} \\
  % Co-authors go here
}

\begin{document}
\maketitle

\begin{abstract}
Tool-calling is necessary but insufficient for biological research engineering.
Agents can shell out to bioinformatics CLIs, but the surface they land on is
hostile: stateful pipelines, undocumented config conventions, and silent
failure modes that look like success. We argue that the missing piece is a
\emph{harness}: opinionated geometric primitives, reproducibility scaffolding,
and agent-friendly APIs over normally-unfriendly biological surfaces. We
demonstrate this with a worked variant effect prediction (VEP) example driven
end-to-end from a natural-language prompt, using the \texttt{manylatents-omics}
extension to the \texttt{manylatents} framework. The harness, the experiment,
and this paper were produced by an agent operating against the same repository
the reader can clone today. % TODO(agent): tighten after results land.
\end{abstract}

\section{Introduction}
\label{sec:intro}

% TODO(agent): expand stakes — current pipelines for VEP require manual model
% selection, config wiring, and CSV plumbing. The agent failure mode is silent
% misalignment: predictions look reasonable but the wrong embedding space was
% used.

\section{Background}
\label{sec:background}

% TODO(agent): brief on (a) Evo2 / Orthrus / ESM3 as foundation models,
% (b) cross-modal embedding alignment, (c) why VEP is the right testbed.

\section{Method}
\label{sec:method}

We use the \texttt{manylatents-omics} extension as the substrate. Configs in
\texttt{experiments/configs/manylatents-omics/experiment/} act as prompts:
\texttt{encode\_evo2.yaml} for DNA encoding, \texttt{encode\_esm3.yaml} for
protein encoding, and \texttt{alignment\_matrix.yaml} for cross-modal $k$-NN
Jaccard overlap.

% TODO(agent): describe the alignment matrix computation. Reference §3.2 of
% the manylatents-omics ARCHITECTURE.md for the formal definition.

\section{Results}
\label{sec:results}

% TODO(agent): produce Table 1 from results/alignment.csv after running:
%   cd experiments/tools/manylatents-omics
%   .venv/bin/python -m manylatents.main +experiment=alignment_matrix
% Then generate Figure 1 via experiments/analysis/NN_figures.py::figure_alignment.

\section{Discussion}
\label{sec:discussion}

% TODO(agent): one paragraph each on (a) what the harness made cheap, (b) what
% the agent still needed human input for, (c) the failure modes encountered.

\section{Limitations}
\label{sec:limitations}

% TODO(agent): be honest about what the demo did and did not cover.

\section*{Acknowledgments}

This paper, the experiments cited within, and the harness used to produce
both, were authored in collaboration with a Claude Code agent operating
against the public \texttt{lrw-vep-ub2026} repository.

\bibliographystyle{plainnat}
\bibliography{../shared/bib/references}

\end{document}
